\documentclass[a4paper,onesided,12pt]{report}
%\usepackage{helvet}
%\renewcommand{\familydefault}{\sfdefault}
\usepackage{styles/fbe_tez}
\usepackage[utf8x]{inputenc} % To use Unicode (e.g. Turkish) characters
\renewcommand{\labelenumi}{(\roman{enumi})}
\usepackage{amsmath, amsthm, amssymb}
 % Some extra symbols
\usepackage[bottom]{footmisc}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{mathtools}
\graphicspath{{images/}} % Graphics will be here

\usepackage{multirow}
\usepackage{subfigure}
\usepackage{algorithm}
\usepackage{algorithmic}
%\pagestyle{empty}
%\includeonly{introduction} % To only process the given file


% For highligts:
\usepackage{xcolor} % For highligts
\usepackage{soul}
\usepackage{pdfpages}  % For model pdf
\usepackage{pdflscape} % for landscape
\usepackage{microtype} % to hide warnings abput width
\usepackage{cleveref} % for chapter refs


%\newtheorem{thm}{Theorem}[chapter]
%\newtheorem{prop}[thm]{Proposition}
%\newtheorem{lem}[thm]{Lemma}
%\newtheorem{cor}[thm]{Corollary}
% COVER PAGE
\title{INVESTIGATION OF THE METABOLIC CHANGES IN THE ADAPTIVE EVOLUTION OF S. CEREVISIAE STRAINS BY USING GENOME-SCALE METABOLIC MODELS}
\turkcebaslik{S. CEREVISIAE SUŞLARININ ADAPTİF EVRİMİNDEKİ METABOLİK DEĞİŞİKLİKLERİN GENOM ÖLÇEKLİ METABOLİK MODELLERLE İNCELENMESİ}
\degree{B.S., Molecular Biology and Genetics, \\
Canakkale Onsekiz Mart University, 2018}
\author{Handan Çetin}
\program{Computational Science and Engineering}
\subyear{2021}

% APPROVED BY PAGE
\cosuperi{Prof. Ş. Kutlu Ülgen}
\cosuperii{Prof. Z. Petek Çakar}
\examineri{Assoc. Prof. Arzucan Özgür}
\examinerii{Assist. Prof. Betül Uralcan}
\examineriii{Assist. Prof. Pınar Pir}
\dateofapproval{30.07.2021}

\begin{document}

\pagenumbering{roman}
\makemstitle % M.S. thesis
\makeapprovalpage
\begin{acknowledgements}
  \input{chapters/acknowledgements.tex}
\end{acknowledgements}
\begin{abstract}
  \input{chapters/abstract.tex}
\end{abstract}
\begin{ozet}
  \input{chapters/abstract_tr.tex}
\end{ozet}
\tableofcontents
\listoffigures
\listoftables
\begin{symbols}
% The title will be typeset as "LIST OF SYMBOLS".
%
% Use a separate \sym command for each symbols definition.
% First, Latin symbols in alphabetical order
\sym{$h$}{hour}
\sym{S}{Stoichiometric coefficient matrix}
\sym{v}{Flux vector}
\sym{}{}
\sym{$\alpha$}{Alpha}
\sym{$\beta$}{Beta}
\sym{$\Delta$}{Delta}
\sym{}{}

\end{symbols}

\begin{abbreviations}
 % Abbreviations in alphabetical order
\sym{ALE}{Adaptive Laboratory Evolution}
\sym{DIY}{Do-It-Yourself}
\sym{EMS}{Ethyl Methane Sulfonate}
\sym{FBA}{Flux Balance Analysis}
\sym{FVA}{Flux Variability Analysis}
\sym{GPR}{Gene-Protein-Reaction}
\sym{GSMM}{Genome-Scale Metabolic Model}
\sym{GUR}{Glucose Uptake Rate}
\sym{HRB}{Hit-and-Run on a Box }
\sym{LP}{Linear Programming}
\sym{MR}{Metabolic Regulation}
\sym{MOMA}{Minimization of Metabolic Adjustment}
\sym{PhPP}{Phenotype Phase Plane}
\sym{PR}{Post-Transcriptional Regulation}
\sym{QP}{Quadratic Programming}
\sym{TR}{Transcriptional Regulation}
\sym{YMM}{Yeast Minimal Media}
\sym{YPD}{Yeast Extract-Peptone-Dextrose}
\end{abbreviations}


\pagenumbering{arabic}
\input{chapters/introduction.tex}
\input{chapters/methods.tex}
\input{chapters/results.tex}
\input{chapters/discussion.tex}
\input{chapters/conclusion.tex}

% \nocite{NewEntry1,NewEntry2,NewEntry3,NewEntry4,NewEntry5,
% NewEntry6,
% NewEntry7,NewEntry8,NewEntry9,NewEntry10,NewEntry11,NewEntry12}

%\cite{*}
\bibliographystyle{styles/fbe_tez_v11}
\bibliography{references}

\appendix

\chapter[RESULTS OF DIFFERENTIAL EXPRESSION ANALYSIS \\
	\hspace*{2.95cm} AND INTEGRATION INFORMATION]{RESULTS OF DIFFERENTIAL EXPRESSION ANALYSIS AND INTEGRATION INFORMATION}
The content is given in the CD.

\chapter[RECONSTRUCTED MODEL FILES]{RECONSTRUCTED MODEL FILES}
The content is given in the CD.

\chapter[SOLUTION VECTORS OF ANALYSES]{SOLUTION VECTORS OF ANALYSES}
The content is given in the CD.


\end{document}
